### 2019-12-20 article-information-2019.tex
proofing

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[information,article,submit,moreauthors,pdftex]{definitions/mdpi}

% If you would like to post an early version of this manuscript  a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,3,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger}$}

% Authors, for metadata in PDF
\AuthorNames{Navdeep Gill, Patrick Hall, Kim Montgomery, and Nicholas Schmidt}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad H2O.ai\\
$^{2}$ \quad BLDS, LLC\\
$^{3}$ \quad George Washington University}

% Contact information of the corresponding author
\corres{Correspondence: phall@h2o.ai; nschmidt@bldsllc.com}

% Current address and/or shared authorship
\firstnote{All authors contributed equally to this work.} 
%\secondnote{}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools. The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on publicly available mortgage data. For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance. The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents: adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures. By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}

% Keywords
\keyword{Machine Learning; Neural Network; Gradient Boosting Machine; Interpretable; Explanation; Fairness; Disparate Impact; Python}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} 
%% Remove this when starting to work on the template.
%\section{How to Use this Template}
%The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e., ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate. ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.  ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}. While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been resolved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}

Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an appealable decision-making framework, tests for DI and other social discrimination, and public and open source software resources. In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models. Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results. Section \ref{sec:con} closes this manuscript with a brief summary of the outlined methods, materials, results, and discussion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}\label{sec:m_and_m}

Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:

\begin{itemize}[leftmargin=*,labelsep=5.8mm]
	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
	\item \textbf{Discrimination testing methods}: AIR, ME, SMD and confusion matrix metrics -- \textsection\ref{ssec:di}
	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
\end{itemize}

To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data. The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}. The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g. \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g. \cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), can be difficult to understand, trust, and appeal, even after the application of post-hoc explanation techniques, explanation analysis and discrimination testing are applied only to the constrained interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}. Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.

Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a template workflow for future users of similar methods and tools. Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discussed discrimination testing methods include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Notation}\label{ssec:not}

To facilitate descriptions of data and modeling, explanatory, and discrimination testing techniques, notation for input and output spaces, datasets, and models is defined.

\subsubsection{Spaces}

	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
		\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.  An arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
		\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
		\item Learned output responses of models are contained in the set $\mathcal{\hat{Y}}$. % For regression models, the set $\mathcal{\hat{Y}}_r$ is also contained in a $C$-dimensional output space, $\mathcal{\hat{Y}}_r \subset \mathbb{R}^{C_r}$. For classification models, the set $\mathcal{\hat{Y}}_c$ typically contains a column vector for each unique class in $\mathcal{Y}$. Hence, $\mathcal{\hat{Y}}_c$ is contained in a $C'$-dimensional output space,  $\mathcal{\hat{Y}}_c \subset \mathbb{R}^{C'_c}$.
	\end{itemize}

\subsubsection{Datasets}

	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
		\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
		\item Each $i$-th observation of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
		$[x_0^{(i)}, x_1^{(i)}, \dots, x_{\textit{P}-1}^{(i)}]$, with corresponding $i$-th labels in $\mathbf{Y}, \mathbf{y}^{(i)}$, and corresponding predictions in $\mathbf{\hat{Y}}, \mathbf{\hat{y}}^{(i)}$. % = [y_0^{(i)}, y_1^{(i)}, \dots, y_{\textit{C}-1}^{(i)}]$.
		\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observations: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$. %\\ $\mathbf{x}^{(i)} \in \mathcal{X}$, $\mathbf{y}^{(i)} \in \mathcal{Y}$.
		\item Each $j$-th input column vector of $\mathbf{X}$ is denoted as $X_j = [x_{j}^{(0)}, x_{j}^{(1)}, \dots, x_{j}^{(N-1)}]^T$.
	\end{itemize}

\subsubsection{Models}

	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
		\item A type of ML model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$: 
		$ \mathbf{X}, \mathbf{Y} \xrightarrow{\mathcal{A}} g$, such that $g \approx f$.
		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
		\item The model to be explained and tested for discrimination is denoted as $g$.
	\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulated Data}\label{ssec:sim_data}

Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:

\begin{equation}
\begin{aligned}\label{eq:sim}\small
f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
\end{aligned}
\end{equation}

\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$. In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies. In this manuscript, the signal generating function and input features are modified in several ways.  Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings. 
%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so. For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.  Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing. The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows. The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.  Within the training set, a 5 fold cross validation indicator is used for training all models. For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mortgage Data}\label{ssec:mort_data}

The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis. This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public. Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.  The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, loan to value (LTV) ratio, debt to income (DTI) ratio, and other important financial descriptors. The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age. Because the data includes information on these protected class characteristics, certain metrics that can be indicative of discrimination in lending can be calculated directly using the HDMA data.

The mortgage dataset analyzed here is a random sample of consumer-anonymized loans from the HDMA database. These loans are a subset of all originated mortgage loans in the 2018 HMDA data that were chosen to represent a relatively comparable group of consumer mortgages.  A selection of features is used to predict whether a loan is \textit{high-priced}, i.e., the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.  After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contains ten input features and the binary outcome, \textit{high-priced}. The data is split into a training set with 160,338 loans and a marker for 5 fold cross validation and a test set containing 39,662 loans. While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.  Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Monotonic Gradient Boosting Machine}\label{ssec:mgbm}

MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each $X_j$ and $\mathbf{y}$ independently. The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.

\begin{equation}
\begin{aligned}\label{eq:gbm}
g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
\end{aligned}
\end{equation}

\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$. For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:

\begin{equation}
\begin{aligned}
\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{MGBM}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
\end{aligned}
\end{equation}

\noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights. Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$. For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction. For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to some $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$. The following rules and constraints ensure positive monotonicity in $\Theta_b$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$.

\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered. 
\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the associated $\theta_{b,j,k-1}$ set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} < \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $ w_{b,j,k,L} \le w_{b,j,k,R}$ and $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
\end{enumerate}

\noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems. Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$. Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$. For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched. Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$. Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.

Herein, two $g^\text{MGBM}$ models are trained.  One on the simulated data and one on the mortgage data. In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment. For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Subsection \ref{ssec:soft}.

\subsection{Explainable Neural Network}\label{ssec:xnn}

XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}. XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.  Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions. XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum_j\beta_{k,j}x_j$.

\begin{equation}
\begin{aligned}
\label{eq:xnn}
g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J=\mathcal{P}-1}_{j=0}\beta_{k, j}x_j)
\end{aligned}
\end{equation}

\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:

\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, which should learn higher magnitude weights for important $X_j$, is known as the \textit{projection layer}. It is fully connected to each input $X_j$. Each hidden unit in the projection layer may optionally include a bias term.
\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}. Each $n_k$ is a neural network, which can be parameterized to suit a given modeling task. To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$. Each $n_k$ can contain several bias terms.
\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$. Again, subnetwork output is restricted to 1-dimension for interpretation and visualization purposes.
\end{enumerate}

Here, each $g^\text{XNN}$ is trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data. Each $g^\text{XNN}$ is assessed in five training folds and in a test data partition. $L_1$ regularization is applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof. The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size. Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times. For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}

PD plots are a widely-used method for describing and plotting the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}. ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}. PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.

Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered. $\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged. $\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched. PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}. Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone. PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shapley Values}\label{ssec:shap}

Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}. Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values should never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}. For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:

\begin{equation}
\label{eq:shap_additive}
\begin{aligned}
g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j \mathbf{z}_j
\end{aligned}
\end{equation}

\noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness. Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$. Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.

\begin{equation}
\label{eq:shap_contrib}
\begin{aligned}
\phi_{j} = \sum_{S \subseteq \mathcal{P} \setminus \{j\}}\frac{|S|!(\mathcal{P} -|S| -1)!}{\mathcal{P}!}[g_x(S \cup \{j\}) - g_x(S)]
\end{aligned}
\end{equation}

\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5. Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some set $\mathbf{X}$.

Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$. Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}. Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Sections \ref{sec:m_and_m} and \ref{sec:res}. Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}. Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Discrimination Testing Metrics}\label{ssec:di}

Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in industries like Financial Services. The discussed metrics are also representative of fair lending analyses and pair well with the mortgage data. Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI). DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.  With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.  In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.  However, other factors, such as hyperparameter or algorithm choices, can drive DI.  Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.

The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings. One is known as marginal effects (ME). ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.

\begin{equation}
\label{eq:me}
\begin{aligned}
\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1))
\end{aligned}
\end{equation}

\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute $X_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group,  and $\text{Pr}(\hat{\mathbf{y}}|X_j)$ indicates a conditional probability regarding model outcomes, $\hat{\mathbf{y}}$, and some feature, $X_j$, evaluated over the relevant partition of data.  ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.

\begin{equation}
\label{eq:air}
\begin{aligned}
\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)}
\end{aligned}
\end{equation}

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome.  Another long-standing measure of DI is standardized mean difference (SMD). SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending. It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}. SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

\begin{equation}
\label{eq:smd}
\begin{aligned}
\text{SMD} \equiv \frac{\bar{\hat{\mathbf{y}}}_p - \bar{\hat{\mathbf{y}}}_c}{ \sigma_{\hat{\mathbf{y}}}}
\end{aligned}
\end{equation}

\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.  Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.  In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.  Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.

A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.  Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.  It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model. Moreover, practical significance is also needed by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production. Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.  One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.  ``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.

The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.  The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v. Partida}.\footnote{\textit{Castaneda v. Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v. United States.}\footnote{\textit{Hazelwood School Dist. v. United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Resources}\label{ssec:soft}

Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:res}

Results are laid out for the simulated and mortgage datasets. Accuracy is compared for unconstrained, less interpretable $g^{\text{GBM}}$ and $g^{\text{ANN}}$ models and constrained, more interpretable $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models. Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretability, post-hoc explanation, and discrimination testing results are presented.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulated Data Results}

Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}. For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance. For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer. Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$. Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.

\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_sim}

Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.

\begin{table}[H]
\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
\centering
\begin{tabular}{ccccc}
\toprule
\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
\midrule
$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
\bottomrule
\label{tab:sim_acc}
\end{tabular}
\end{table}

\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model.

\subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}

Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}. Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
\label{fig:sim_mgbm_glob}
\end{figure}

\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.

Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}. PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$. ICE serves several purposes. It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is also a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior. PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain. Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction. Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
\label{fig:sim_mgbm_glob_pdp_ice}
\end{figure}

$X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE. Also, as the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$. Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is probably not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}. The one-dimensional monotonic constraints may hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data. This is an interesting result for responsible ML practitioners. In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.

Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}. These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}. Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/sim_mgbm_loc.png}
\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
\label{fig:sim_mgbm_loc}
\end{figure}

\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result. For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative. At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive. In each case, large local contributions generally follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.

Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique. Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$. Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_xnn_glob.png}
\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
\label{fig:sim_xnn_glob}
\end{figure}

\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features. The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures. Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.

Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$. \ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights. The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms. Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.

\begin{figure}[H]
\centering
\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
\label{fig:sim_xnn_glob_ridge}
\end{figure}

\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks. Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units. In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.  Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.  Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio. As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.

Local Deep SHAP feature importance in Figure \ref{fig:sim_xnn_loc} supplements the global interpretability of $g^\text{XNN}$ displayed in Figures \ref{fig:sim_xnn_glob} and \ref{fig:sim_xnn_glob_ridge}. Local Deep SHAP values are extracted from the projection layer of $g^\text{XNN}$ and reported in the probability space. Deep SHAP values can be calculated for any arbitrary $g^\text{XNN}(\mathbf{x})$, allowing for detailed, local summarization of individual model predictions.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/sim_xnn_loc.png}
\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
\label{fig:sim_xnn_loc}
\end{figure}

\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x}^{(i)})$ predictions.

\subsubsection{Discrimination Testing Results}\label{ssec:dis_sim}

Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.  Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The selection of which cutoff to use in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.  For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.  For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score. But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.  Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.} 
Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities. Table \ref{tab:sim_disc} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs. 1.06).  However, in Table \ref{tab:sim_n_acc} the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs. 0.401). This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$. This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.  Of course, the choice of which model is truly fairer is a policy question.

\begin{table}[H]
\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
\centering
\begin{tabular}{cccccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy}  								& \textbf{FPR}									& \textbf{FNR} \\
\midrule
Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.168\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\	
\midrule			 
Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.163\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
\midrule
Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.186\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\	
\midrule
Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.138\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\		
\bottomrule
\label{tab:sim_n_acc}
\end{tabular}
\end{table}

For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.  Of note is that 12.0\% is not a meaningful difference without context.  If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 12.0\% could represent an extremely large difference and would require remediation.  But if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.  As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 12.0\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population. Protected group 1’s AIR under $g^\text{XNN}$ is 0.727, below the EEOC 4/5ths rule threshold.  It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}). Together these would indicate that there may be evidence of illegal DI.  As with ME and other measures, the reasonableness of this disparity is not clear outside of context.  However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review. Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.

\begin{table}[H]
\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
\centering
\begin{tabular}{cccccccc}
\toprule
\textbf{Model}		& \textbf{Protected Class}										& \textbf{Control Class}								& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD}   								& \textbf{FPR Ratio}							& \textbf{FNR Ratio} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.16 \\ 1.86\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
\midrule 
$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.03 \\ 1.35\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\										
\bottomrule
\label{tab:sim_disc}
\end{tabular}
\end{table}

\noindent SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.  (This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs. -0.274), but both are close to Cohen’s small effect threshold of -0.20. Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mortgage Data Results}

Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort}.  $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants. Assuming that small fit differences on static test data do not outweigh the need for intrinsic model interpretability and reliable post-hoc explainability in high-stakes, human-centered, or regulated applications, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.

\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec:c_v_uc_mort}

Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data. This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g., numerical encoding of categorical features and missing values. This preprocessing appears to hamstring some of the tree-based models' inherent capabilities. $g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).

\begin{table}[H]
\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccccc}
\toprule
\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
\midrule	
$g^\text{GBM}$		& 0.795	& 0.828	& 0.252	& 0.276 \\
$g^\text{MGBM}$		& 0.765	& 0.814	& 0.259	& 0.278 \\
$g^\text{ANN}$		& 0.865	& 0.871	& 0.231	& 0.262 \\
$g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
\bottomrule
\label{tab:mort_acc}
\end{tabular}
\end{table}

\noindent Regardless of the fit differences between the two families of hypothesis models, the difference between the fit of constrained and unconstrained variants within the two types of models is small for the GBMs and negligible for ANNs, 3\% and < 1\% worse fit respectively, averaged across the metrics reported in Table \ref{tab:mort_acc}.

\subsubsection{Interpretability and Post-hoc Explanation Results}

Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}. $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/mort_mgbm_glob.png}
\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_mgbm_glob}
\end{figure}

\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.

Domain knowledge was used to positively constrain DTI ratio and LTV ratio and to negatively constrain income and the loan term flag under $g^\text{MGBM}$. The monotonicity constraints for DTI ratio and LTV ratio are confirmed for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data in Figure \ref{fig:mort_mgbm_glob_pdp_ice}. Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles for ICE and on average with PD. Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$. Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and perhaps due to less noise in the mortgage data, this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data. DTI ratio and LTV ratio also appear to have sparse regions in their univariate distributions. The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}$ appears to carry reasonable predictions learned from populous domains into the sparse domains of both features.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/mort_mgbm_glob_pdp_ice.png}
\caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
\label{fig:mort_mgbm_glob_pdp_ice}
\end{figure}

Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value. Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves. While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$, and lower,  exhibit different prediction behavior with respect to property value. Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior. Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn. For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions. However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and appear in a sparse region of property value. This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}). However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.

In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data. The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/mort_mgbm_loc.png}
\caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.}
\label{fig:mort_mgbm_loc}
\end{figure}

\noindent Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to \textit{accurately} summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual. In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.

Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}$ on the mortgage test data. $g^\text{XNN}$ distributes importance more evenly across business drivers and puts stronger emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$. Like $g^\text{MGBM}$, $g^\text{XNN}$ puts little emphasis on the other flag features.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/mort_xnn_glob.png}
\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_xnn_glob}
\end{figure}

\noindent As compared to $g^\text{MGBM}$, $g^\text{XNN}$ assigns higher importance to property value, loan amount, and income, and lower importance on LTV ratio and DTI ratio.

The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}. \ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ have large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, have small or near-zero magnitude weights. Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B} and color-coded to pair with their corresponding $\gamma_k$ weight. As in the subsection \ref{ssec:int_ex_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}. In both the simulated and mortgage data, $g^\text{XNN}$ $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.

\begin{figure}[H]
\centering
\includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
\label{fig:mort_xnn_glob_ridge}
\end{figure}

\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function. For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.

To compliment the global interpretability of $g^\text{XNN}$, Figure \ref{fig:mort_xnn_loc} displays local Shapley values for selected individuals, estimated from the projection layer using Deep SHAP in the $g^\text{XNN}$ probability space.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/mort_xnn_loc.png}
\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_xnn_loc}
\end{figure}

\noindent While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc} their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions, even for neural network architectures.

\subsubsection{Discrimination Testing Results}\label{ssec:dis_mort}

Tables \ref{tab:mort_n_acc} and \ref{tab:mort_disc} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males. As with the simulated data, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the FPRs, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).  Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.  Thus, one might consider FPR ratios as a measure of the class-control disparities.  FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs. 2.10) in Table \ref{tab:mort_disc},  but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs. 0.315) in Table \ref{tab:mort_n_acc}.  This is the same pattern seen in the simulated data results in Section \ref{ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.  Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.

\begin{table}[H]
\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{ccccccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy} 									& \textbf{FPR}											& \textbf{FNR} \\
\midrule
Black 		& 2,608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} 			& \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\	
\midrule			 
White 		& 28,361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}			& \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\		
\midrule
Female 		& 8,301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular}		& \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}			& \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
\midrule
Male 		& 13,166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}			& \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
\bottomrule
\label{tab:mort_n_acc}
\end{tabular}
\end{table}

Both ME and AIR show higher disparities for blacks under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$.  Blacks receive high-priced loans 21.4\% more frequently using $g^{\text{XNN}}$ vs. 18.3\% for $g^{\text{MGBM}}$.  Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ show AIRs that are statistically significantly below parity (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), and which are also below the EEOC’s 0.80 threshold. This would typically indicate cause for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section \ref{ssec:disc_rem}. On the other hand, women improve under $g^{\text{XNN}}$ vs. $g^{\text{MGBM}}$ (MEs of 3.6\% vs. 4.1\%; AIRs of 0.955 vs. 0.948). The AIRs, while statistically significantly below parity, are well above the EEOC’s threshold of 0.80.  In most situations, the values of these metrics alone would not likely flag a model for further review, though there may be other considerations which would lead to concern.

\begin{table}[H]
\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{cccccccc}
\toprule
\textbf{Model}	& \textbf{Protected Class}	& \textbf{Control Class}	& \textbf{AIR}	& \textbf{ME}	& \textbf{SMD}  & \textbf{FPR Ratio}		& \textbf{FNR Ratio} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular}	& \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular}	& \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
\midrule 
$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\							
\bottomrule
\label{tab:mort_disc}
\end{tabular}
\end{table}

Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.  These exceed Cohen’s guidelines of 0.5 for a medium effect size and would likely trigger further review.  Female SMDs are well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$, respectively.  Similar to results for female AIR, these values alone are unlikely to prompt further review.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:disc}

\subsection{The Burgeoning Python Ecosystem for Responsible Machine Learning}

MGBM and XNN interpretable model architectures were selected for this text because they are straightforward variants of popular unconstrained ML models. If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models. The same can be said of the presented explanation methods and discrimination tests. Due to their post-hoc nature, they can often be shoe-horned into existing ML work flows and pipelines. While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML. Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/blueprint.png}
\caption{A diagram of a proposed holistic ML work flow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, and other review and appeal mechanisms can create an understandable and trustworthy ML system.}
\label{fig:blueprint}
\end{figure}

Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell. In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages. Among several types of currently available interpretable models, practitioners can now use Python to evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{See: \href{https://github.com/xiyanghu/OSDT}{Optimal sparse decision trees}.}\textsuperscript{,}\footnote{See: \href{https://github.com/cfchen-duke/ProtoPNet}{\textit{This looks like that} interpretable deep learning}.} Additional, relevant references and Python functionality include:

\begin{itemize}[leftmargin=*,labelsep=5.8mm]

\item \textbf{Exploratory data analysis (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.

\item \textbf{Sparse feature extraction}: \texttt{H2OGeneralizedLowRankEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}{\texttt{h2o}} \cite{udell2016generalized}.

\item \textbf{Privacy preprocessing and private models}: differential privacy and private models in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} and \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{diffprivlib}, \cite{ji2014differential}, \cite{pate}, \cite{abadi2016deep}.

\item \textbf{Post-hoc explanation}: structured data explanations with \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}}, image classification explanations with \href{https://github.com/marcoancona/DeepExplain}{\texttt{DeepExplain}}, and natural language explanations with \href{https://github.com/allenai/allennlp}{\texttt{allennlp}} \cite{wachter2017counterfactual}, \cite{grad_attr}, \cite{wallace2019allennlp}.

\item \textbf{Discrimination testing}: with \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.

\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification with \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} \cite{kamiran2012data}, \cite{zhang2018mitigating}, \cite{lfr}, \cite{kamiran2012decision}.

\item \textbf{Model debugging}: with \href{https://github.com/bethgelab/foolbox}{\texttt{foolbox}}, \href{https://github.com/SALib/SALib}{\texttt{SALib}}, \href{https://github.com/tensorflow/cleverhans}{\texttt{tensorflow/cleverhans}}, and \href{https://github.com/tensorflow/model-analysis}{\texttt{tensorflow/model-analysis}} \cite{rauber2017foolbox}, \cite{papernot2018cleverhans}, \cite{modeltracker}, \cite{papernot2018marauder}.

\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.,  \href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, \href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.

\end{itemize}

\noindent See: \href{https://github.com/jphall663/awesome-machine-learning-interpretability}{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.

\subsection{Appeal and Override of Automated Decisions}

Interpretable model architectures and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions. As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain. Post-hoc explanations generally help describe and summarize mechanisms and decisions, potentially yielding an even clearer understanding of ML models. Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}. Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}. Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, providing people affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention. ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions is difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.

\subsection{Impact of Discrimination Testing on Model Use and Adoption}

\subsection{Viable Discrimination Remediation Approaches}\label{ssec:disc_rem}

\subsection{Intersectional and Non-static Problems in Machine Learning}

The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own. However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}. High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}. This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{\textcolor{red}{24}}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems. Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena. Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML. However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination. Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory. Accuracy, discrimination, and security characteristics of a system can change over time as well. Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems. Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:con}

This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods. These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible ML work flow for high-stakes, human-centered, or regulated ML applications. The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects. Of course, there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems. While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work (and many, many others) illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{NG, data cleaning, GBM and MGBM assessment and results; PH, primary author; KM, ANN and XNN implementation, assessment, and results; NS, secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{This work received no external funding.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{Wen Phan for work in formalizing notation. Sue Shay for editing. Andrew Burt for ideas around the transparency paradox.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank. Wells Fargo is a customer of, and investor in, H2O.ai and a customer of BLDS, LLC. However, communications regarding XNN between Wells Fargo and the authors have been extremely limited prior to and during the drafting of this manuscript. Moreover, Wells Fargo exerted absolutely no editorial control over the text or results herein.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this text: 
AI -- artificial intelligence, 
AIR - adverse impact ratio,
ALE - accumulated local effect,
ANN -- artificial neural network, 
APR -- annual percentage rate,
AUC -- area under the curve,
CNN -- convolutional neural network,
CFPB -- Consumer Financial Protection Bureau,
DI -- disparate impact, 
DT -- disparate treatment,
DTI -- debt to income,
EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e. variants GAMs that consider two-way interactions and may incorporate boosting into training, 
EEOC -- Equal Employment Opportunity Commission,
ECOA - Equal Credit Opportunity Act,
EDA -- exploratory data analysis,
EU -- European Union,
FCRA -- Fair Credit Reporting Act,
FNR -- false negative rate,
FPR -- false positive rate,
GAM -- generalized additive model, 
GBM -- gradient boosting machine, 
GDPR - General Data Protection Regulation,
HMDA -- Home Mortgage Disclosure Act,
ICE -- individual conditional expectation, 
LTV -- loan to value,
ME -- marginal effect,
MGBM -- monotonic gradient boosting machine, 
ML -- machine learning, 
PD -- partial dependence, 
RMSE -- root mean square error,
SGD -- stochastic gradient descent, 
SHAP -- Shapley additive explanation,
SMD - standardized mean difference,
SR -- supervision and regulation, 
US -- United States, 
XNN -- explainable neural network.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

%=====================================
% References, variant A: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{article-information-2019}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
%\begin{quote}
%This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
%\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Subsection}
%\unskip
%\subsubsection{Subsubsection}

%Bulleted lists look like this:
%\begin{itemize}[leftmargin=*,labelsep=5.8mm]
%\item	First bullet
%\item	Second bullet
%\item	Third bullet
%\end{itemize}

%Numbered lists can be added as follows:
%\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
%\item	First item 
%\item	Second item
%\item	Third item
%\end{enumerate}

%The text continues here.

%\subsection{Figures, Tables and Schemes}

%All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

%\begin{figure}[H]
%\centering
%\includegraphics[width=2 cm]{Definitions/logo-mdpi}
%\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
%\end{figure}

%Text

%Text

%\begin{table}[H]
%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
%\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
%\begin{tabular}{ccc}
%\toprule
%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
%\midrule
%entry 1		& data			& data\\
%entry 2		& data			& data\\
%\bottomrule
%\end{tabular}
%\end{table}

%Text

%Text

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\textwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\textwidth}{1pt}
%\end{listing}

%\subsection{Formatting of Mathematical Components}

%This is an example of an equation:

%\begin{equation}
%a + b = c
%\end{equation}
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

%Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
%\begin{Theorem}
%Example text of a theorem.
%\end{Theorem}

%The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
%\begin{proof}[Proof of Theorem 1]
%Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
%\end{proof}
%The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}

%Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Materials and Methods}

%Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

%Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusions}

%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Patents}
%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%\noindent 
%\begin{tabular}{@{}ll}
%MDPI & Multidisciplinary Digital Publishing Institute\\
%DOAJ & Directory of open access journals\\
%TLA & Three letter acronym\\
%LD & linear dichroism
%\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendix
%\section{}
%\unskip
%\subsection{}
%The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations an

%=====================================
% References, variant B: internal bibliography
%=====================================
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; %Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{Samples of the compounds ...... are available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}